---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Welcome! Let's get you started with bbsBayes.

First we'll make sure we have the right software installed, we'll fetch the 
breeding bird data, and then we'll run through some example workflows.



## Install [cmdstanr](https://mc-stan.org/cmdstanr/)

Because bbsBayes uses Stan to run the Bayesian models, we need to make
sure we have cmdstanr and cmdstan both installed.

It is recommended to run this is a fresh R session or restart your current
session
```{r, eval = FALSE}
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", 
                                       getOption("repos")))
```

Now we should be able to use cmdstanr to install cmdstan
```{r, eval = FALSE}
cmdstanr::install_cmdstan()
```

Let's check that everything went as planned

```{r}
cmdstanr::cmdstan_path()
cmdstanr::cmdstan_version()
```


> Problems? Check out cmdstanr's vignette on 
> [Getting Started](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)



## Download BBS data

Now we'll fetch the BBS data using functions included in bbsBayes.

> Note that most bbsBayes functionality can be explored **without** downloading
> BBS data by using the included sample data. Specify `sample_data = TRUE`
> in the first `stratify()` step (see the next section). 

First we'll load the package.
```{r}
library(bbsBayes)
```

There are four types of BBS data that can be downloaded:
- Two levels `state` and `stop` (only `state` works with bbsBayes models)
- Two releases `2020` and `2022`

Unless you have a specific reason to need a different version, the default 
(level `state` and release `2022`) is likely the one you want. However
you can download all data sets and specify which one you wish to use in the
`stratify()` step.

```{r, eval = FALSE}
fetch_bbs_data()                  # Default
fetch_bbs_data(release = "2020")  # Specify a different release
```


```{r, echo = FALSE, eval = !have_bbs_data()}
fetch_bbs_data()
```

## Workflow

Now that you have cmdstanr installed and the BBS data downloaded, we'll walk
through a general workflow for modelling species trends with bbsBayes.

### Stratify the data

The first step in any bbsBayes analysis is to stratify the data. 
In this step we choose a stratification type as well as a species to explore. 

```{r}
s <- stratify(by = "bbs_usgs", species = "Snowy Owl")
```

We can also play around with the included sample data (Pacific Wrens)
```{r}
s <- stratify(by = "bbs_cws", sample_data = TRUE) # Only Pacific Wren
```

Stratifications included in bbsBayes are 
`r paste0(names(bbs_strata), collapse = ",")`. See the articles on
[stratifications](articles/stratification) and 
[custom stratifications](articles/custom_stratification) for more details and 
examples.

Available species can be explore using the `search_species()` function.

### Prepare the data

Once we have stratified the data, we can now prepare it for use in a model. 
In this step data will be filtered to omit routes with too few samples, etc.
See `prepare_data()` for more details on how you can customize this step.

```{r}
p <- prepare_data(s)
```

### Prepare the model

Next we will prepare the model parameters and initialization values.
See `prepare_model()` for more details on how you can customize this step.

```{r}
md <- prepare_model(p, model = "first_diff")
```


### Run model

Now we can run the model.

The default `iter_sampling` and `iter_warmup` are 1000 and the default `chains`
is 3. In the interest of speed for this example, we are using much lower values,
but note that this almost certainly will result in problems with our model.

```{r, eval = FALSE}
m <- run_model(md, iter_sampling = 20, iter_warmup = 20, chains = 2)
```

```{r, echo = FALSE}
m <- pacific_wren_model
```

### Explore the model outputs

First we will generate indices according to different regional categories.
By default these are `continent` and `stratum`. 

```{r}
i <- generate_indices(model_output = m)
```

We can visually explore these indices by accessing the `indices` item in the list

```{r}
i[["indices"]]
```

We can also plot these indicies. 

```{r}
p <- plot_indices(indices = i, add_observed_means = TRUE)
```

Note that we get one plot for each region and regional category, in this case
that means one plot for the continent, and one for each stratum.

```{r}
names(p)
```

We can plot them individually by pulling a plot out of the list

```{r}
p[["continent"]]
p[["CA_AB_10"]]
```

Next we can calculate the specific trends based on these indices

```{r}
t <- generate_trends(i)
```

We can plot these on a map

```{r}
plot_map(t)
```

```{r, include = FALSE}
unlink(list.files(pattern = "^BBS_STAN_(\\w|\\d|-)+.(csv|rds)"))
```

