---
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Install [cmdstanr](https://mc-stan.org/cmdstanr/)
It is recommended to run this is a fresh R session or restart your current
session
```{r, eval = FALSE}
install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", 
                                       getOption("repos")))
```

## Install cmdstan

- Problems? Check out cmdstanr's vignette on 
[Getting Started](https://mc-stan.org/cmdstanr/articles/cmdstanr.html)

```{r, eval = FALSE}
cmdstanr::install_cmdstan()
```


## Check cmdstanr installation

- Useful to include in model runs for reproducibility

```{r}
cmdstanr::cmdstan_path()
cmdstanr::cmdstan_version()
```


```{r}
library(bbsBayes)
```


## Download BBS data

- Two levels `state` and `stop` (only `state` works with bbsBayes models)
- Two releases `2020` and `2022`

```{r, eval = FALSE}
fetch_bbs_data(force = TRUE) # Defaults to 2022
fetch_bbs_data(release = "2020", force = TRUE)
```


```{r, echo = FALSE, eval = !have_bbs_data()}
fetch_bbs_data()
```

## Stratify the data

```{r}
s <- stratify(by = "bbs_usgs", species = "Snowy Owl")
s <- stratify(by = "bbs_usgs", sample_data = TRUE) # Only Pacific Wren
```

## Prepare the data

```{r}
p <- prepare_data(s, min_max_route_years = 2)
```

## Run model

- Super fast, just for illustration

```{r, eval = FALSE}
m <- run_model(p, model = "slope",
               iter_sampling = 10, iter_warmup = 10, chains = 2)
```

## Look at the model outputs

```{r, eval = FALSE}
i <- generate_indices(model_output = m,
                            regions = c("continental",
                                        "national",
                                        "prov_state",
                                        "stratum"))

tp <- plot_indices(indices = i, add_observed_means = TRUE)
```



## Troubleshooting


```
Warning message:
In fun(libname, pkgname) :
rgeos: versions of GEOS runtime 3.10.2-CAPI-1.16.0
and GEOS at installation 3.8.0-CAPI-1.13.1differ
```

```{r, eval = FALSE}
pak::pkg_remove("rgeos")
pak::pkg_install("rgeos")
```




