---
title: "Stratification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stratification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette we'll explore the various ways you can stratify your data in 
preparation for running the models. 

You can use existing, pre-defined strata, subset an existing strata, or create
your own custom stratification, either using a completely new set of spatial
data, or by modifying the spatial polygons of an existing strata.

Let's get started!


First, we'll load the packages we need.
```{r setup}
library(bbsBayes)
library(sf)       # Spatial data manipulations
library(dplyr)    # General data manipulations
library(ggplot2)  # Plotting
```

Then we'll make sure we have the BBS data downloaded
```{r}
have_bbs_data()
```

If not, install with `fetch_bbs_data()`
```{r, eval = !have_bbs_data()}
fetch_bbs_data()
```

## Stratifying with existing categories

The existing stratifications are `r paste0(names(bbs_strata), collapse = ", ")`

You can take a look at how these stratify your data by looking at the maps 
included in bbsBayes with `load_map()`.

```{r}
ggplot(data = load_map("bbs_cws"), aes(fill = strata_name)) +
  geom_sf() +
  scale_fill_viridis_d(guide = NULL)
```


To stratify BBS data, you can use these existing stratifications by 
specifying `by = "name"` in the `stratify()` function.

```{r}
s <- stratify(by = "bbs_usgs", species = "Canada Jay")
```


## Subsetting an existing stratification

But what if you want to use the BBS CWS stratification, but only really want
to look at Canadian regions?

In this case you'll subset the BBS CWS stratification and give `stratify()` that
data set in addition to the specification. 

In addition to maps, stratifications are available as data frames in the 
`bbs_strata` object.

```{r}
names(bbs_strata)
bbs_strata[["bbs_cws"]]
```

We can now modify and use this data frame as we like.
```{r}
my_cws <- filter(bbs_strata[["bbs_cws"]], country == "United States of America")
s <- stratify(by = "bbs_cws", species = "Canada Jay", strata_custom = my_cws)
```

Note that the stratification is now "bbs_cws" and "subset"

```{r}
s$meta_data
```


## Custom stratification - New map

To define a completely different stratification, you'll need to provide a 
spatial data object with polygons defining your strata.

In our example we'll use WBPHS stratum boundaries for 2019.
This is available from available from the US Fish and Wildlife Service Catalogue: <https://ecos.fws.gov/ServCat/Reference/Profile/142628>

You can either download it by hand, or with the following code.
```{r}
t <- tempfile()

download.file(url = "https://ecos.fws.gov/ServCat/DownloadFile/213149",
              destfile = t)
unzip(t)
unlink(t)
```

To use this file in bbsBayes, we need to load it as an sf object using the sf 
package.
```{r}
map <- sf::read_sf("WBPHS_stratum_boundaries_2019.shp")
map
ggplot(map, aes(fill = factor(STRAT))) + 
  geom_sf() +
  scale_fill_viridis_d(guide = NULL)

```

We see that it has one column that reflects the stratum names. First we'll
rename this column to `strata_name` which is what `stratify()` requires.

```{r}
map <- rename(map, strata_name = STRAT)
map
```


Now we have the spatial data and relevant information to pass to `stratify()`.

When using a custom stratification, the `by` argument becomes the name you want
to apply. Let's use something informative, but short (although there's no limit).
We also need to give the function our map.

```{r}
s <- stratify(by = "WBPHS_2019", species = "Canada Jay", strata_custom = map)
```

> Note that strata names are automatically put into lower case for consistency. 

We can take a quick look at the output, by looking at the meta data and routes
contained therein.
```{r}
s[["meta_data"]]
s[["routes_strata"]]
```

To get a different look we can also plot this data on top of our map using
ggplot2. Note that we use `factor()` to ensure the stratum names are 
categorical.

```{r}
r <- s[["routes_strata"]] %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

ggplot() +
  geom_sf(data = map, aes(fill = factor(strata_name)), alpha = 0.3) +
  geom_sf(data = r, aes(colour = factor(strata_name)), size = 1) +
  scale_fill_viridis_d(aesthetics = c("colour", "fill"), guide = NULL)

```

Based on the message we received during stratification and this map, 
it looks as if our custom stratification is excluding some BBS data. 

We can re-run the stratification with `return_omitted = TRUE` which will attach
a data frame of omitted strata to the output.
```{r}
s <- stratify(by = "WBPHS_2019", species = "Canada Jay", strata_custom = map,
              return_omitted = TRUE)
s[["routes_omitted"]]
```

Let's take a look at this visually.

```{r}
omitted <- st_as_sf(s[["routes_omitted"]], coords = c("longitude", "latitude"), 
                    crs= 4326)

ggplot() +
  geom_sf(data = map, aes(fill = factor(strata_name)), alpha = 0.3) +
  geom_sf(data = r, aes(colour = factor(strata_name)), size = 1, alpha = 0.5) +
  geom_sf(data = omitted, size = 0.75, alpha = 0.5) +
  scale_fill_viridis_d(aesthetics = c("colour", "fill"), guide = NULL) 
```


We could also look at this visually by comparing the stratified data to the raw
BBS data.

First we load the raw routes data, then we'll add it to our plot.

This plot may take a little while to run!

```{r}
bbs_routes <- load_bbs_data()[["routes"]] %>%
  st_as_sf(coords = c("longitude", "latitude"), crs= 4326)

ggplot() +
  geom_sf(data = map, aes(fill = factor(strata_name)), alpha = 0.3) +
  geom_sf(data = r, aes(colour = factor(strata_name)), size = 1) +
  geom_sf(data = bbs_routes, size = 0.5) +
  scale_fill_viridis_d(aesthetics = c("colour", "fill"), guide = NULL) 
```

So either way, we can clearly see, spatially, which points have been omitted
from our data.

## Custom stratification - Modifying existing BBS maps

Stratify by custom stratification, using sf map object
e.g, with east/west divide of southern Canada with BBS CWS strata

First we'll start with the CWS BBS data
```{r}
map <- load_map("bbs_cws")
```

We'll modify this by first looking only at provinces (omitting the northern territories), transforming to the GPS CRS (4326), and ensuring the resulting 
polygons are valid.
```{r}
new_map <- map %>%
  filter(country_code == "CA", !prov_state %in% c("NT", "NU", "YT")) %>%
  st_transform(4326) %>%
  st_make_valid()
```

Now we can crop this map to make a western and an eastern portion, defined by 
longitude and latitude (which is why we first transformed to the GPS CRS).
```{r}
west <- st_crop(new_map, xmin = -140, ymin = 42, xmax = -95, ymax = 68) %>%
  mutate(strata_name = "west")
east <- st_crop(new_map, xmin = -95, ymin = 42, xmax = -52, ymax = 68) %>%
  mutate(strata_name = "east")
```

Now we'll bind these together and transform back to the original CRS

```{r}
new_strata <- bind_rows(west, east) %>%
  st_transform(st_crs(map))

ggplot() +
  geom_sf(data = map) +
  geom_sf(data = new_strata, aes(fill = strata_name), alpha = 1)
```


Looks good! Let's use it in our stratification and take a look at the points 
afterwards to ensure they've been categorized appropriately.

```{r}
s <- stratify(by = "canada_ew", species = "Canada Jay", 
              strata_custom = new_strata)

s$meta_data
routes <- s$routes_strata %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

ggplot() +
  geom_sf(data = new_strata, aes(fill = strata_name), alpha = 1) +
  geom_sf(data = routes, aes(shape = strata_name))

```

Now we're good to continue on with `prepare_data()`...






