---
title: "Assessing Model Convergence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assessing Model Convergence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bbsBayes)
library(ggplot2)
```

First make sure you have the BBS data downloaded
```{r}
have_bbs_data()
```

If not, install with `fetch_bbs_data()`
```{r, eval = !have_bbs_data()}
fetch_bbs_data()
```

Let's start by running a quick (and dirty) model looking at the Pacific Wren.

```{r, eval = FALSE}
m <- stratify(by = "bbs_cws", sample_data = TRUE) %>%
  prepare_data() %>%
  prepare_model(model = "first_diff") %>%
  run_model(iter_sampling = 20, iter_warmup = 20, chains = 2)

```

Or we can use the example model included in bbsBayes, `pacific_wren_model`
```{r}
m <- pacific_wren_model
```


Now we can calculate our convergence metrics
```{r, warning=FALSE}
conv <- get_convergence(m)
conv
```

Wow, there are a *lot* of variables here.

We can visualize this by transforming the data frame and using ggplot2

```{r}
tconv <- tidyr::pivot_longer(conv, cols = c(ess_bulk, ess_tail, rhat))

ggplot(data = tconv, aes(x = variable, y = value, colour = variable_type)) +
    geom_point() +
    facet_wrap(~name, scales = "free_y") +
    scale_colour_viridis_d(guide = FALSE)
```

We can also choose to extract only some variables. To see which ones are available,
use the `get_model_vars()` function.

```{r}
get_model_vars(m)
```


We can also extract summary information from the model via the helper function `get_summary()` (wrapper for  `cmdstanr::summary()`)

```{r, warning=FALSE}
get_summary(m)
```


