---
title: "Custom Stratifications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Stratifications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The Stratification vignette covers how to stratify your data using existing or 
custom stratifications. Here we will explore how to use custom stratifications 
in later steps, such as `generate_indices()`.

First, we'll load the packages we need.
```{r setup}
library(bbsBayes)
library(sf)       # Spatial data manipulations
library(dplyr)    # General data manipulations
library(ggplot2)  # Plotting
```


Then make sure you have the BBS data downloaded
```{r}
have_bbs_data()
```

If not, install with `fetch_bbs_data()`
```{r, eval = !have_bbs_data()}
fetch_bbs_data()
```

We'll start with a custom stratification example we used in the original
Stratification vignette (see that vignette for more details regarding the
`stratify()` step)

First we need a spatial data object with polygons defining the strata.

In our example we'll use WBPHS stratum boundaries for 2019.
This is available from available from the US Fish and Wildlife Service Catalogue: <https://ecos.fws.gov/ServCat/Reference/Profile/142628>

You can either download it by hand, or with the following code.
```{r, eval = !file.exists("WBPHS_stratum_boundaries_2019.zip")}
z <- "WBPHS_stratum_boundaries_2019.zip"

download.file(url = "https://ecos.fws.gov/ServCat/DownloadFile/213149",
              destfile = z)
unzip(z)
unlink(z)
```

We'll load and prepare this data.
```{r}
map <- sf::read_sf("WBPHS_stratum_boundaries_2019.shp") %>%
  rename(strata_name = STRAT)
```

And stratify our data!
```{r}
s <- stratify(by = "WBPHS_2019", species = "Canada Jay", strata_custom = map)
```

Once stratified, the next couple of steps are the same, regardless of whether 
we're using a custom stratification or not.
```{r}
p <- prepare_data(s)
```

Prepare and run the model (note this is not an appropriate model, as we're aiming for speed
rather than accuracy).
```{r}
m <- prepare_model(p, model = "first_diff") %>%
  run_model(chains = 2, iter_sampling = 20, iter_warmup = 20)
```

Now we can start to look at the indices and trends related to our model.

`generate_indices()` creates indices at different regional levels. By default
these are "continent" and "stratum", but you can add "prov_state", "bcr",
"bcr_by_country" (where appropriate). You can also create your own regional
divisions and provide them as a `regions_index` data frame.

For example, let's imagine we would like to calculate regional indices for 
each stratum, country, province/state, as well as for a custom division of 
east vs. west. 

First we'll need to tell the function which strata belong to which province or 
state, as well as which belong to the 'east' and which to the 'west.

For the first, we can use a helper function `assign_prov_state()`. This function
takes your map of strata and assigns each strata to a province or state
depending on the amount of overlap. By default it will warn if the amount of
overlap is less than 75%, but in this case, we will lower that cutoff to 60%.
The plot gives us a chance to make a quick assessment of whether we're happy 
with how the various strata have been assigned.

```{r}
rindex <- assign_prov_state(map, min_overlap = 0.6, plot = TRUE)
```

Next we'll define the east/west divide by hand. If we plot the strata by 
name, we can pick out which are eastern and which western.

```{r}
ggplot(rindex) +
  geom_sf(data = load_map(type = "North America")) +
  geom_sf() +
  geom_sf_text(aes(label = strata_name))

rindex <- mutate(rindex, 
                 east_west = if_else(as.numeric(strata_name) < 50 | 
                                       as.numeric(strata_name) > 74,
                                     "west", "east"))

ggplot(data = rindex) +
  geom_sf(data = load_map(type = "North America")) +
  geom_sf(data = rindex, aes(fill = east_west), alpha = 0.5)

```

Now we'll create our indices and trends
```{r}
i <- generate_indices(
  m, 
  regions = c("stratum", "country", "prov_state", "east_west"),
  regions_index = rindex)

t <- generate_trends(i)
```

We can plot each of these indices with `plot_indices()` which creates a list
of plots. 

```{r}
#| fig-asp: 0.5
#| fig-width: 10
p <- plot_indices(i)
names(p)

library(patchwork)
p[["east"]] + p[["west"]]
```

Finally we can even create geofaceted plots (which is only possible in our case
because we assigned our strata to Provinces and States and calculated indices
for these regions).
```{r}
plot_geofacet(i, trends = t, multiple = TRUE)
```

```{r, include = FALSE}
unlink(list.files(pattern = "^BBS_STAN_(\\w|\\d|-)+.(csv|rds)"))
```
